---
title: "Create a heatmap of the USA"
output: github_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "imgs/heatmap_map",
  out.width = "100%"
)
```

There are multiple methods of creating maps in R if you have longitude and latitude coordinates or full state or country counts. This tutorial will assume you have some familiarity to tidyverse and magrittr style R programs. For a full discussion on the benefits of tidyverse-style R programming, see [Wickham et al, 2019](https://tidyverse.tidyverse.org/articles/paper.html)

## R Libraries required

You might need to install the packages if they are not already installed.

```{r dep, eval = FALSE}
install.packages(c("tidyverse", "magrittr", "ggmap", "RColorBrewer", "usmap"))
```

Then load the libraries into R.

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(magrittr)
library(usmap)
#library(ggmap)         # <= requires a google key
#library(RColorBrewer)
#library(sf)            # <= eventually use sf methods
```

## Input File

We are going to assume you have a count of items by state. Please download the input file from this [link]().



```{r}
# === Load the Excel Spreadsheet of state and counts
data <- readxl::read_excel("data/usa.xlsx")
head(data)
```

Now let's combine that with states.

```{r}

plot_usmap(include = c(data$TwoLetters)) +
  labs(title = "The United States",
       subtitle = "These are the 50 states.")

```
