---
title: "Create a heatmap of the USA"
output: github_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "imgs/heatmap_map",
  out.width = "100%"
)
```

There are multiple methods of creating maps in R if you have longitude and latitude coordinates or full state or country counts. This tutorial will assume you have some familiarity to [tidyverse](https://tidyverse.tidyverse.org/) and [magrittr](https://cran.r-project.org/web/packages/magrittr/vignettes/magrittr.html) style R programs. For a full discussion on the benefits of tidyverse-style R programming, see ["Welcome to the Tidyverse" (Wickham et al, 2019)](https://tidyverse.tidyverse.org/articles/paper.html)

## R Libraries required

You might need to the following R libraries (or packages) if they are not already installed.

```{r dep, eval = FALSE}
install.packages(c("tidyverse", "magrittr", "usmap"))
```

Then load the libraries into R.

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(magrittr)
library(usmap)
#library(ggmap)         # <= requires a google key
#library(RColorBrewer)
#library(sf)            # <= eventually use sf methods
```

## Input File

We are going to assume you have a count of items by state. Please download the input file (`usa.xlsx`) from this [link](data/usa.xlsx) and place it into a `data` folder.



```{r}
# === Load the Excel Spreadsheet of state and counts
data <- readxl::read_excel("data/usa.xlsx")
head(data)
```

Now let's combine that with states.

```{r}

plot_usmap(include = c(data$TwoLetters)) +
  labs(title = "The United States",
       subtitle = "Just the outline")

```
